---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

# call data
```{r}
flights_joined <- read_csv(here::here("clean_data/flights_joined.csv")) %>% 
  clean_names()
```


# HypotHesis test: EWR & LGA

**H0:** The time average of departure delay of Newark airport is the same of the La Guardian airport.
$$
H_0: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} - \mu_{\ Guardian \ aiport \ time \ avg \ departure \ delay} = 0
$$ 

**H1:** The time average of departure delay of Newark airport is higher than the La Guardian airport.
$$
H_1: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} -\mu_{\ Guardian \ aiport \ time \ avg \ departure \ delay} > 0
$$


```{r}
# ----------------------------------------------------------------------------
# decide significance level (a)
# ----------------------------------------------------------------------------
a <- 0.05


# ----------------------------------------------------------------------------
# null sample distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_LGA <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "LGA")) %>%
  specify(dep_delay ~ origin) %>% #(response = dep_delay, explanatory = origin)
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("EWR", "LGA"))

# ----------------------------------------------------------------------------
# Calculate observed statistic
# ----------------------------------------------------------------------------
obs_statistic_EWR_LGA <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "LGA")) %>%
  specify(dep_delay ~ origin) %>%
  calculate(stat = "diff in means", order = c("EWR", "LGA"))
obs_statistic_EWR_LGA

# ----------------------------------------------------------------------------
# Calculate p-value
# ----------------------------------------------------------------------------
null_distribution_EWR_LGA %>%
  get_p_value(direction = "right", obs_stat = obs_statistic_EWR_LGA)

# ----------------------------------------------------------------------------
# Visualize the p-value on the null distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_LGA %>%
  visualise() +
  shade_p_value(direction = "right", obs_stat = obs_statistic_EWR_LGA) + 
  theme_minimal() +
  labs(title = "Visualize the p-value on the null distribution") +
  scale_x_continuous(breaks = seq(-3, 10, 1)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```
##conclusion
p-value  > α (1 > 0.05): there is not a strong enough evidence to reject H0. 
We do not have enough evidence that the time average of departure delay in
Newark is higher than in the La Guardian airport.




# HypotHesis test: EWR & JFK


**H0:** The time average of departure delay of Newark airport is the same of the John F Kennedy International Airport.
$$
H_0: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} - \mu_{\ Jonh.F \ Kennedy \  aiport \ time \ avg \ departure \ delay} = 0
$$ 

**H1:** The time average of departure delay of Newark airport is higher than the John F Kennedy International Airport.
$$
H_1: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} -\mu_{\ Jonh.F \ Kennedy \  aiport \ time \ avg \ departure \ delay} > 0
$$

```{r}
# ----------------------------------------------------------------------------
# decide significance level (a)
# ----------------------------------------------------------------------------
a <- 0.05

# ----------------------------------------------------------------------------
# null sample distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_JFK <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "JFK")) %>%
  specify(dep_delay ~ origin) %>% #(response = dep_delay, explanatory = origin)
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("EWR", "JFK"))

# ----------------------------------------------------------------------------
# Calculate observed statistic
# ----------------------------------------------------------------------------
obs_statistic_EWR_JFK <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "JFK")) %>%
  specify(dep_delay ~ origin) %>%
  calculate(stat = "diff in means", order = c("EWR", "JFK"))
obs_statistic_EWR_JFK 

# ----------------------------------------------------------------------------
# Calculate p-value
# ----------------------------------------------------------------------------
null_distribution_EWR_JFK %>%
  get_p_value(direction = "right", obs_stat = obs_statistic_EWR_JFK)

# ----------------------------------------------------------------------------
# Visualize the p-value on the null distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_JFK %>%
  visualise() +
  shade_p_value(direction = "right", obs_stat = obs_statistic_EWR_JFK) + 
  theme_minimal() +
  labs(title = "Visualize the p-value on the null distribution") +
  scale_x_continuous(breaks = seq(-3, 10, 1)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```
##conclusion
p-value  > α (0.493 > 0.05) there is not a strong enough evidence to reject H0.
We do not have enough evidence that the time average of departure delay in 
Newark is higher than in the John F Kennedy International Airport.
