---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

# call data
```{r}
flights_joined <- read_csv(here::here("clean_data/flights_joined.csv")) %>% 
  clean_names()
```


# HypotHesis test: EWR & LGA

**H0:** The time average of departure delay of Newark airport is the same of the La Guardian airport.
$$
H_0: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} - \mu_{\ Guardian \ aiport \ time \ avg \ departure \ delay} = 0
$$ 

**H1:** The time average of departure delay of Newark airport is higher than the La Guardian airport.
$$
H_1: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} -\mu_{\ Guardian \ aiport \ time \ avg \ departure \ delay} < 0
$$


```{r}
# ----------------------------------------------------------------------------
# decide significance level (a)
# ----------------------------------------------------------------------------
a <- 0.05


# ----------------------------------------------------------------------------
# null sample distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_LGA <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "LGA")) %>%
  specify(dep_delay ~ origin) %>% #(response = dep_delay, explanatory = origin)
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("EWR", "LGA"))

# ----------------------------------------------------------------------------
# Calculate observed statistic
# ----------------------------------------------------------------------------
obs_statistic_EWR_LGA <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "LGA")) %>%
  specify(dep_delay ~ origin) %>%
  calculate(stat = "diff in means", order = c("EWR", "LGA"))
obs_statistic_EWR_LGA

# ----------------------------------------------------------------------------
# Calculate p-value
# ----------------------------------------------------------------------------
null_distribution_EWR_LGA %>%
  get_p_value(direction = "right", obs_stat = obs_statistic_EWR_LGA)

# ----------------------------------------------------------------------------
# Visualize the p-value on the null distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_LGA %>%
  visualise() +
  shade_p_value(direction = "right", obs_stat = obs_statistic_EWR_LGA) + 
  theme_minimal() +
  labs(title = "Visualize the p-value on the null distribution") +
  scale_x_continuous(breaks = seq(-3, 10, 1)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```
##conclusion
The p-value is lower than the significance level (0.05), so there is strong 
evidence to not reject H0 in favour of H1 and conclude that the time average of 
delay of departure in Newark airport is not higher than the La Guardian airport.

There are 5% of probability to reject the hypothesis that the difference of 
the means of the time of departure delay is the same in both airports, being
this hypothesis true.




# HypotHesis test: EWR & JFK


**H0:** The time average of departure delay of Newark airport is the same of the John F Kennedy International Airport.
$$
H_0: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} - \mu_{\ Jonh.F \ Kennedy \  aiport \ time \ avg \ departure \ delay} = 0
$$ 

**H1:** The time average of departure delay of Newark airport is higher than the John F Kennedy International Airport.
$$
H_1: \mu_{\ Newark \ aiport \ time \ avg \ departure \ delay} -\mu_{\ Jonh.F \ Kennedy \  aiport \ time \ avg \ departure \ delay} > 0
$$

```{r}
# ----------------------------------------------------------------------------
# decide significance level (a)
# ----------------------------------------------------------------------------
a <- 0.05

# ----------------------------------------------------------------------------
# null sample distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_JFK <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "JFK")) %>%
  specify(dep_delay ~ origin) %>% #(response = dep_delay, explanatory = origin)
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("EWR", "JFK"))

# ----------------------------------------------------------------------------
# Calculate observed statistic
# ----------------------------------------------------------------------------
obs_statistic_EWR_JFK <- flights_joined %>%
  filter(dep_delay > 0) %>% 
  filter(origin %in% c("EWR", "JFK")) %>%
  specify(dep_delay ~ origin) %>%
  calculate(stat = "diff in means", order = c("EWR", "JFK"))
obs_statistic_EWR_JFK 

# ----------------------------------------------------------------------------
# Calculate p-value
# ----------------------------------------------------------------------------
null_distribution_EWR_JFK %>%
  get_p_value(direction = "right", obs_stat = obs_statistic_EWR_JFK)

# ----------------------------------------------------------------------------
# Visualize the p-value on the null distribution
# ----------------------------------------------------------------------------
null_distribution_EWR_JFK %>%
  visualise() +
  shade_p_value(direction = "right", obs_stat = obs_statistic_EWR_JFK) + 
  theme_minimal() +
  labs(title = "Visualize the p-value on the null distribution") +
  scale_x_continuous(breaks = seq(-3, 10, 1)) +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```
##conclusion
p-value is is greater than the significance level (0.05), then we lack sufficient
evidence to reject H0 and so we fail to reject H0. Based on our sample, we do
not have enough evidence that the time average of departure delay in Newark 
Airport is higher than in the John F Kennedy International Airport.