---
title: "R Notebook"
output: html_notebook
---

# Decision tree
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(GGally)
library(rpart)
library(rpart.plot)
library(modelr)
library(caret)
library(tsibble)
library(ranger)
```


# call data
```{r}
flights_joined <- read_csv(here::here("clean_data/flights_joined.csv")) %>% 
  clean_names() 
```
```{r}
flights_joined <- flights_joined %>% 
  filter(origin == "EWR") %>% 
  select(dep_delay, wind_dir, wind_speed, wind_gust, visib)  
```

```{r}
flights_joined <- flights_joined %>% 
  filter(origin == "EWR") %>% 
   # select(dep_delay, wind_dir, wind_speed, wind_gust, visib) 
  select(dep_delay, c(temp:visib)) %>% 
  select(-wind_gust, -dewp)
```


# Correlation between weather variables
```{r}
flights_joined %>% 
  ggcorr(label = TRUE)
```
Temp and dewp has a strong correlation, and dewp has a strong correlation with
humid. Will be removed dewp variable.

Wind speed and wind gust has a perfect correlation. Will be removed the wind gust.


## clean data
```{r}
tree_weather <- flights_joined %>% 
  mutate(dep_delay = as.factor(if_else(condition = dep_delay <= 0,
                                       true = "OnTime",
                                       false = "Delay"))) %>% 
  drop_na()


  # mutate(month = month(time_hour),
  #        week   = week(time_hour)) %>%
  # group_by(month, week) %>%
  # mutate(across(where(is.numeric), ~ coalesce(.x, mean(.x, na.rm = TRUE)))) %>%
  # ungroup() %>%
  # drop_na() %>% 
  # select(-time_hour,- month, -week)


# tree_weather %>% 
#   distinct(visib)

```

```{r}
tree_weather <- flights_joined %>% 
  filter(origin == "EWR") %>% 
  select(-time_hour, -origin) %>% 
  mutate(dep_delay = as.factor(if_else(condition = dep_delay <= 0,
                                       true = "OnTime",
                                       false = "Delay"))) %>%
  # mutate(month = month(time_hour),
  #         week   = week(time_hour)) %>%
  # #  group_by(month, week) %>%
  # mutate(across(where(is.numeric), ~ coalesce(.x, median(.x, na.rm = TRUE)))) %>%
  # 
  # ungroup() %>%
  drop_na()
# select(-time_hour,- month, -week)
```


## test/ train & check balanced
```{r}
# ------------------------------------
# Testing and training dataset
# ------------------------------------
n_data <- nrow(tree_weather)

# create a test sample index
test_index <- sample(1:n_data, size = n_data*0.2)

# create test set
tree_weather_test  <- slice(tree_weather, test_index)

# create training set
tree_weather_train <- slice(tree_weather, -test_index)


# ------------------------------------
# check balanced sets
# ------------------------------------
tibble(
  tree_weather_test %>%
    janitor::tabyl(dep_delay)
  )


tibble(
  tree_weather_train %>%
    janitor::tabyl(dep_delay)
  )


# ------------------------------------
# Create a decision tree
# ------------------------------------
tree_weather_fit <- rpart(
  formula = dep_delay ~ ., 
  data = tree_weather_train, 
  method = 'class'
)

rpart.plot(tree_weather_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 4)
```


```{r}
rpart.rules(tree_weather_fit, cover = TRUE)
```



```{r}
# ------------------------------------
# add the predictions
# ------------------------------------
tree_test_pred <- tree_weather_test %>%
  add_predictions(tree_weather_fit, type = 'class') %>% 
  relocate(pred, .after = dep_delay)

tree_test_pred
```

## confusion matrix
```{r}


confusionMatrix(
  tree_test_pred$pred, 
  tree_test_pred$dep_delay
  )
```
The main diagonal represents correctly-predicted values, with the top right 
values showing false positives and the bottom left being false negatives. 

The more accurate the decision tree, the higher the main diagonal values will be. 

Accuracy represents the probability of our prediction being correct.


# Random florest 
```{r}
rf_classifier <- ranger(dep_delay ~ ., 
                        data = tree_weather_train, 
                        importance = "impurity", 
                        num.trees = 1000, 
                        mtry = 2, 
                        min.node.size = 5)

rf_classifier


# ------------------------------
# Importance
# -----------------------------
importance(rf_classifier)
```



```{r}
# ----------------------------
# add predictors
# ----------------------------
rf_test_pred <- tree_weather_test %>%
  mutate(pred = predict(rf_classifier, data = tree_weather_test)$predictions, 
         .after = "dep_delay")

rf_test_pred
```

```{r}
confusionMatrix(rf_test_pred$pred, rf_test_pred$dep_delay)
```

